---
title: "eUSR"
author: "Yuning Li, Siming Su"
date: "9/27/2022"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Library setup 
```{r}
library(statnet)
library(ergm)
library(igraph)
library(dplyr)
library(network)
library(ggplot2)
```


## Import data 
```{r}
week6_online <- read.csv("week6_on_cleaned.csv")
week9_online <- read.csv("week9_on_cleaned.csv")

week6_online <- week6_online[,-1]
```



## grade shift
```{r}
grade <- c('A', 'A-', 'B+', 'B', 'B-', 'C+', 'C', 'C-', 'F')
gpa <- c(4, 3.7, 3.3, 3.0, 2.7, 2.3, 2.0, 1.7, 0)

week6_online$mid_gpa <- 0
for (i in 1:nrow(week6_online)){
  week6_online$mid_gpa[i] <- gpa[match(week6_online$midterm_grade[i], grade)]
}

week9_online$final_gpa <- 0
for (i in 1:nrow(week9_online)){
  week9_online$final_gpa[i] <- gpa[match(week9_online$final_grade[i], grade)]
}

common_id <- intersect(week6_online$Id, week9_online$Id)

week9_online$grade_shift <- 2
for (i in 1:nrow(week9_online)){
  if (week9_online$Id[i] %in% common_id){
    mid_gpa <- week6_online$mid_gpa[which(week6_online$Id == week9_online$Id[i])]
    if (week9_online$final_gpa[i] > mid_gpa){
      week9_online$grade_shift[i] <- 1
    } else {
      week9_online$grade_shift[i] <- 0
    }
  }
}
week9_online$grade_shift <- as.factor(week9_online$grade_shift)
```

## attitude of hw
Easy < Just right < moderate difficulty < difficult < very difficult (1-5)
```{r}
week6_online$attitude_hw <- as.factor(week6_online$attitude_hw)
```

## pstat dept
```{r}
pstat_dpt <- c('Actuarial Science', 'Financial Mathematics and Statistics', 
                'Statistics and Data science')

week6_online$stat <- 2
for (i in 1:nrow(week6_online)){
  if (week6_online$first_major[i] %in% pstat_dpt | week6_online$second_major[i] %in% pstat_dpt){
    week6_online$stat[i] <- 1
  } else{
    week6_online$stat[i] <- 0
  }
}

week6_online$stat <- as.factor(week6_online$stat)

week9_online$stat <- 2
for (i in 1:nrow(week9_online)){
  if (week9_online$first_major[i] %in% pstat_dpt | week9_online$second_major[i] %in% pstat_dpt){
    week9_online$stat[i] <- 1
  } else{
    week9_online$stat[i] <- 0
  }
}

week9_online$stat <- as.factor(week9_online$stat)
```

## indegree of perception as strong course material
```{r}
week6_online$indegree <- 0

for (i in 1:nrow(week6_online)){
  count <- 0
  id <- week6_online$Id[i]
  for (j in 10:15){
    for (z in 1:nrow(week6_online)){
      if (week6_online[z,j] == id){
        count <- count + 1
      }
    }
  }
  week6_online$indegree[i] <- count
}


week9_online$indegree <- 0

for (i in 1:nrow(week9_online)){
  count <- 0
  id <- week9_online$Id[i]
  for (j in 9:14){
    for (z in 1:nrow(week9_online)){
      if (week9_online[z,j] == id){
        count <- count + 1
      }
    }
  }
  week9_online$indegree[i] <- count
}


## indegree shift
week9_online$indegree_shift <- 400
for (i in 1:nrow(week9_online)){
  if (week9_online$Id[i] %in% common_id){
    mid_indegree <- week6_online$indegree[which(week6_online$Id == week9_online$Id[i])]
    week9_online$indegree_shift[i] <- week9_online$indegree[i] - mid_indegree
  }
}


```

## trimmed dataset
```{r}
week9_online_sub = week9_online[week9_online$Id %in% common_id,]
week6_online_sub = week6_online[week6_online$Id %in% common_id,]

write.csv(week9_online_sub,"/Users/simingsu/Documents/summer_research/social_networking/social_network_analysis/week9_online.csv", row.names = FALSE)
write.csv(week6_online_sub,"/Users/simingsu/Documents/summer_research/social_networking/social_network_analysis/week6_online.csv", row.names = FALSE)
```

## convert to igraph
```{r}
inter_on6 = read.csv("total_edge_interaction_week6_on.csv")
strong_on6 = read.csv("total_edge_strong_week6_on.csv")
inter_on9 = read.csv("total_edge_interaction_week9_on.csv")
strong_on9 = read.csv("total_edge_strong_week9_on.csv")

inter_on6 = inter_on6[,2:3]
strong_on6 = strong_on6[,2:3]
inter_on9 = inter_on9[,2:3]
strong_on9 = strong_on9[,2:3]

igraph_on6 = graph.edgelist(as.matrix(inter_on6) ,directed=FALSE)
sgraph_on6 = graph.edgelist(as.matrix(strong_on6) ,directed=FALSE)
igraph_on9 = graph.edgelist(as.matrix(inter_on9) ,directed=FALSE)
sgraph_on9 = graph.edgelist(as.matrix(strong_on9) ,directed=FALSE)

igraph_on6 = igraph_on6 %>% simplify()
sgraph_on6 = sgraph_on6 %>% simplify()
igraph_on9 = igraph_on9 %>% simplify()
sgraph_on9 = sgraph_on9 %>% simplify()

V(igraph_on6)$name = V(igraph_on6)
```


## basic informaiton
```{r}
table(week6_online_sub$gender)
table(week9_online_sub$gender)

## some changes from week 6 to week9 
gender_xy = week6_online_sub %>% inner_join(week9_online_sub, by = "Id")%>% 
  select(c(Id, gender.x, gender.y)) 
gender_xy[gender_xy$gender.x != gender_xy$gender.y, ]

## week 6 reponse rate  
nrow(week6_online)/287
nrow(week9_online)/287

## density comparison (interaction increase !)
edge_density(igraph_on6)
edge_density(igraph_on9)

diameter(igraph_on6)
diameter(igraph_on9)
```


## data management
```{r}
deg_df6 = data.frame(Id = 1:287,
           degree = degree(igraph_on6))
deg_df9 = data.frame(Id = 1:287,
           degree = degree(igraph_on9))

week9_online_sub = week9_online_sub %>% inner_join(deg_df9, by = "Id")
week6_online_sub = week6_online_sub %>% inner_join(deg_df6, by = "Id")
```


### hypo test
```{r}
## gender diff (no difference between gender)
gender_df6 = week6_online_sub[(week6_online_sub$gender=="Female") | (week6_online_sub$gender=="Male"),]
kruskal.test(degree ~ gender, data = gender_df6)

## gender diff (no diff)
gender_df9 = week9_online_sub[(week9_online_sub$gender=="Female") | (week9_online_sub$gender=="Male"),]
kruskal.test(grade_shift ~ gender, data = gender_df9)

summary(aov(final_gpa ~ gender, data = gender_df9))
summary(aov(mid_gpa ~ gender, data = gender_df6))


## gpa versus pstat department (mid diff, final no)
kruskal.test(mid_gpa ~ stat, data = week6_online_sub)
kruskal.test(final_gpa ~ stat, data = week9_online_sub)


## gpa versus attitude toward assignment (diff all)
df_att = week6_online_sub[, c("Id", "attitude_hw")]
week9_online_sub = week9_online_sub %>% inner_join(df_att, by = "Id")


kruskal.test(mid_gpa ~ attitude_hw, data = week6_online_sub)
kruskal.test(final_gpa ~ attitude_hw, data = week9_online_sub)


# visualization
ggplot(week6_online_sub, aes(x=attitude_hw, y=mid_gpa)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)

ggplot(week9_online_sub, aes(x=attitude_hw, y=final_gpa)) + 
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)


## gpa versus indegree (diff all) (strong indegree -> better grade)
cor.test(week6_online_sub$mid_gpa, week6_online_sub$indegree, method = "pearson")
cor.test(week9_online_sub$final_gpa, week9_online_sub$indegree, method = "pearson")
```


## comparison
```{r}
## difference between degree !!!
t.test(week6_online_sub$degree, week9_online_sub$degree, paired = TRUE, alternative = "less")

t.test(week6_online_sub$mid_gpa, week9_online_sub$final_gpa, paired = TRUE, alternative = "less")

## number of triads
## number of mutuals 
## number of transitivity


## closeness week6 vs week9 (nodiff)
close_df6 = data.frame(Id = 1:287,
           closeness = closeness(igraph_on6))
close_df9 = data.frame(Id = 1:287,
           closeness = closeness(igraph_on9))

week9_online_sub = week9_online_sub %>% inner_join(close_df9, by = "Id")
week6_online_sub = week6_online_sub %>% inner_join(close_df6, by = "Id")

week6_online_sub$closeness = week6_online_sub$closeness %>% replace(is.na(.), 0)
week9_online_sub$closeness = week9_online_sub$closeness %>% replace(is.na(.), 0)



## betweenness week6 vs week9(mean diff has a diff)
bet_df6 = data.frame(Id = 1:287,
           between = betweenness(igraph_on6))
bet_df9 = data.frame(Id = 1:287,
           between = betweenness(igraph_on9))

week9_online_sub = week9_online_sub %>% inner_join(bet_df9, by = "Id")
week6_online_sub = week6_online_sub %>% inner_join(bet_df6, by = "Id")

t.test(week6_online_sub$between,
            week9_online_sub$between,
            paired = TRUE)

wilcox.test(week6_online_sub$between,
            week9_online_sub$between,
            paired = TRUE,
            exact = FALSE)

```

## Visualization
```{r}
## draw neighborhood of 14 
## number of interaction
## strong report
## visualization

which.max(degree(igraph_on6))
degree(igraph_on9)


nei_list = neighbors(igraph_on6, 14)
V(igraph_on6)$Id
 V(igraph_on6)[name %in% nei_list]
# select the nodes having these names
selnodes <- V(igraph_on6)[name %in% nei_list]
# get their network neighborhood 
selegoV <- ego(igraph_on6, order=1, nodes = 14, mode = "all", mindist = 0)

# turn the returned list of igraph.vs objects into a graph
selegoG <- induced_subgraph(igraph_on6,unlist(selegoV))

# plot the subgraph
plot(selegoG,vertex.label=V(selegoG)$name)
```



## comparison of two weeks 
```{r}
plot_graph = function(graph){
  l = layout_with_fr(graph)
  l <- norm_coords(l, ymin=-1, ymax=1, xmin=-1, xmax=1) #default -- scaled
  
  V(graph)$size = degree(graph) %>% log()*4
  plot(graph, vertex.label.color="black", 
  vertex.label.cex=.5, vertex.label.dist=0, edge.curved=0.5,layout = l*1.3, rescale= T)
}

plot_graph(igraph_on6)
plot_graph(igraph_on9)

# ## assign colors to different genders
# vcol=V(inter_new)$gender
# vcol[(vcol!="Female")&(vcol!="Male")]="gray50"
# vcol[vcol=="Female"]="tomato"
# vcol[vcol=="Male"]="gold"
# V(inter_new)$color = vcol
# 
# 
# ## assgin the degree size
# V(inter_new)$size = degree(inter_new) %>% log()*4
# 
# plot(inter_new, vertex.label.color="black", 
#      vertex.label.cex=.5, vertex.label.dist=0.5, edge.curved=0.5,layout = l*1.3, rescale= F)
# 
# legend("topleft", legend = c("Female","Male","Other"), pch=21,
#        col=c("tomato","gold","gray50"), pt.bg=c("tomato","gold","gray50"), pt.cex=1, cex=.8, bty="n", ncol=1)
```


## transform it into network object and set attributes
```{r}
## transform it into network
edge_on9<- as.data.frame(get.edgelist(igraph_on9))
edge_on9 <- edge_on9[edge_on9$V1 %in% common_id,]
edge_on9 <- edge_on9[edge_on9$V2 %in% common_id,]
## get to network
net_on9 = network(edge_on9,
                  directed = FALSE,
                  loops = FALSE) 


## get network in week6 for ftiting
edge_on6<- as.data.frame(get.edgelist(igraph_on6))
edge_on6 <- edge_on9[edge_on6$V1 %in% common_id,]
edge_on6 <- edge_on9[edge_on6$V2 %in% common_id,]
## get to network
net_on6 = network(edge_on6,
                  directed = FALSE,
                  loops = FALSE) 




## create attributes dataframe
net_id9 = data.frame(Id = network.vertex.names(net_on9))
df_attr9 = left_join(net_id9, week9_online_sub, by = "Id")

## add midterm grade to df_attr
mid_df = inner_join(net_id9, week6_online_sub, by = "Id")[c("Id","midterm_grade")]
df_attr9 = left_join(df_attr9, mid_df, by = "Id")

df_attr9$grade_shift = as.numeric(df_attr9$grade_shift)
df_attr9$attitude_hw = as.numeric(df_attr9$attitude_hw)


attr_names = c("Id", "final_grade", "gender", "race", "first_major", "class_standing", "grade_shift", "attitude_hw", "indegree", "midterm_grade")

for (name in attr_names){
  network::set.vertex.attribute(net_on9, name, df_attr9[[name]])
}
```


## ERGM
### triying 1st fit
```{r}
fit.01 = ergm(net_on9~ edges + nodefactor('final_grade', base = 2:3) + nodematch('final_grade',diff=T) +
                nodefactor('gender') + nodematch('gender',diff=T) + 
                nodefactor('race') + nodematch('race',diff=T))
summary(fit.01)
```

change of attributes
```{r}
A_grade = c("A-", "A")
B_grade = c("B+", "B-", "B")
C_grade = c("C+", "C-", "C")

net_on9%v%"fgrade_level" = ifelse(net_on9%v%"final_grade" %in% A_grade, "A", 
                                    ifelse(net_on9%v%"final_grade" %in% B_grade, "B",
                                           ifelse(net_on9%v%"final_grade" %in% C_grade, "C", "F")))

net_on9%v%"mgrade_level" = ifelse(net_on9%v%"midterm_grade" %in% A_grade, "A", 
                                    ifelse(net_on9%v%"midterm_grade" %in% B_grade, "B",
                                           ifelse(net_on9%v%"midterm_grade" %in% C_grade, "C", "F")))

net_on9%v%"class_level" = ifelse(net_on9%v%"class_standing" == "Freshman", 0, 
                                    ifelse(net_on9%v%"class_standing" == "Sophomore", 1,
                                           ifelse(net_on9%v%"class_standing" == "Junior", 2,
                                                  ifelse(net_on9%v%"class_standing" == "Senior", 3, 4))))

```

## third fit
```{r}
fit.03 = ergm(net_on9~ edges +      ## number of edges in week9
                nodefactor('final_grade') + nodematch('final_grade', diff = T)+ ## final grade nodefactor and nodematch
                nodefactor('gender') + nodematch('gender',diff=T) +  ## gender nodefactor and nodematch
                nodefactor('race') + nodematch('race',diff=T) +  ## race nodefactor and nodematch
                nodefactor('class_standing')+ nodematch('class_standing',diff=T) +## class standing nodefactor and nodematch
                edgecov(net_on6) + ## week6 network
                nodecov('indegree') + ## perception of peers
                nodefactor("grade_shift") + 
                absdiff("class_level") 
                )

summary(fit.03)
```

```{r}
fit.03.gof <- gof(fit.03)
plot(fit.03.gof)
```


